---
title: "LandTrendr-Paramo - GABAM"
author: "James Millington"
date: "2021-08-03"
output: 
  github_document:
    toc: true
---

Checking [GABAM](https://vapd.gitlab.io/post/gabam/) data compared to our study area 



```{r warning=F, message=F}
library(raster)
library(sf)
library(landscapemetrics)
```


```{r}

NAize <- function(x) { ifelse(x == 0, NA, x) }
gabam <- stack()

years <- 2011:2019
for(yr in years){
  
  r <- raster(paste0('/home/james/wdResearch/LandTrendr-Paramo/Data/GABAM/N10W080_',yr,'.TIF'))
  r <- crop(r, extent(-74, -72, 5, 7))
  r <- calc(r, NAize)
  gabam <- stack(gabam, r)
  
}
names(gabam) <- years
plot(gabam, col='red')


GISpath <- "/home/james/OneDrive/Research/Projects/ColombiaBIO/Fire/Fire GIS Files/" #linux
buf <- st_read(paste0(GISpath,"500mbufferOfComplejos/GroundTruthingBuffer500m.shp"))

#set crs of shp to same as for cci raster
buf <- st_transform(buf, "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

```

```{r}

for(yr in years){
  
  lyr <- paste0('X',yr)
  fires <- lsm_p_area(gabam[[lyr]])
  print(paste0(yr,": ",max(fires['id']), "fires"))
  
}
  
```

There seem to be very few fires... more cloud issues?

> The availability of adequate good quality observations remains another limiting factor for Landsat-based BA detection, since the temporal gaps caused by relatively lower temporal resolution and cloud contamination can result in omission errors, particularly in tropic regions. 

Long _et al._ (2019) https://doi.org/10.3390/rs11050489
