---
title: "LandTrendr-Paramo Explore"
author: "James Millington"
date: "2020-02-04"
output: github_document
---

Email from Kris, 2020-01-30

> I’ve rerun what I think should be our baseline settings with the results downloadable from `Data/URLs/REF1`

> Note there are 6 layers:
>
> 1. Year of Disturbance detection (YOD)
2. Magnitude of disturbance
3. Duration of disturbance
4. Pre-disturbance value
5. Rate of spectral change
6. Disturbance signal-to-noise ratio (DSNR)*

Try loading data (six-layer .tif file)

```{r}
library(raster)

dat <- stack("Data/LTGEE20012020.tif")
names(dat)

names(dat) <- c("Year", "Magnitude", "Duration", "Pre Value", "Spec Chg", "DSNR")

plot(dat)

ncell(dat)
```

Looks like NoData has not been indicated (lots of zeros). Should be fine for first three layers to set NoData = 0. But also, this is a huge raster. So:

- subset to first three layers
- crop to a smaller extent for testing
- set 0 to NA

```{r}

#subset
sdat <- stack(dat[[1]], dat[[2]], dat[[3]])

#cropt
extent(sdat)
crp <- c(-72.3, -72.25, 5.2, 5.25)
sdat <- crop(sdat, crp)

#0 to NA
NAize <- function(x) { ifelse(x == 0, NA, x) }
sdat <- calc(sdat, fun=NAize)

plot(sdat)

```

There's no unique fire ID. Add one using `clump` on `Year` layer:

```{r}

IDs <- clump(sdat[["Year"]])
plot(IDs)
names(IDs) <- "FireID"

sdat <-stack(sdat, IDs)
plot(sdat)

```

Let's check if this worked. If so, all patches should have a SD of zero (because all values within the patch should be identical):

```{r}

tab.years.sd <- zonal(x=sdat[["Year"]], z=sdat[["FireID"]], fun='sd')
tab.years.sd

```

Something strange about FireIDs 3 and 35. Examine those only:

```{r}

NAize3 <- function(x) { ifelse(x != 3, NA, x) }

sdat[["Fire3"]] <- calc(sdat[["FireID"]], NAize3)

freq(sdat[["Fire3"]])

```

This shows `FireID 3` has size one cell, hence `sd = NA`. So this is okay. Let's look at `FireID 35`:

```{r}

NAize35 <- function(x) { ifelse(x != 35, NA, x) }

sdat[["Fire35"]] <- calc(sdat[["FireID"]], NAize35)

dat35.yr <- mask(sdat[["Year"]], sdat[["Fire35"]])
dat35.yr <- trim(dat35.yr)
plot(dat35.yr)

freq(dat35.yr)
unique(dat35.yr)

```

This map shows that cells from two different years are being `clump`ed into the same Fire (FireID == 35). So this approach is not appropriate. 

[This SO post](https://stackoverflow.com/a/30097452) implies we need to loop through Years (classes) creating clumps.

Or could we use the `landscapemetrics` [package](https://r-spatialecology.github.io/landscapemetrics/index.html). Let's try this package first:

```{r}

library(landscapemetrics)

p <- get_patches(sdat[["Year"]])

plot(stack(p))

```

So `get_patches` returns a list of `raster` layers (which can then be `stack`ed)

Let's check the the patch classes identified by `get_patches` matches what we see in the original data:

```{r}

unique(sdat[["Year"]])

```

So now, how can we extract the number of fires in each year? Ideally, we will be able to extract a dataframe year by patch. First, we should [check the landscape](https://r-spatialecology.github.io/landscapemetrics/articles/getstarted.html#checking-your-landscape):

```{r}

check_landscape(sdat[["Year"]])

```

`landscapemetrics` wants us to use CRS with units of m. Check the current CRS: 

```{r}

crs(dat)

```

This is in decimal degrees. Using [this webpage](https://www.latlong.net/lat-long-utm.html) it seems our study area is UTM zone 19N. So re-project to UTM using proj.4 from [EPSG:32619](https://epsg.io/32619):

```{r}

sr <- "+proj=utm +zone=19 +datum=WGS84 +units=m +no_defs" 

sdat <- stack(sdat[["Year"]],sdat[["Magnitude"]],sdat[["Duration"]])
sdat.utm <- projectRaster(sdat, crs = sr, method="ngb")
plot(sdat.utm)

```

And then check landscape again: 
```{r}
check_landscape(sdat.utm)

```

In this landscape check we see `Year` and `Duration` are fine, but `Magnitude` should not be analysed using `landscapemetrics` (due to number of classes, i.e. this is not a categorical variable, but continuous). 

Back to working out how to extract the number of fires in each year:

```{r}
library(tidyverse)

#first calculate patch metric (area)
area <- lsm_p_area(sdat.utm[[1]])
area <- area %>% 
  rename(Year = class, Area_ha = value)
area 

#now we can count fires in each year 
area %>% 
  group_by(Year) %>% 
  tally()

#and plot a bar chart
area %>% 
  group_by(Year) %>% 
  tally() %>%
  ggplot(aes(Year)) +
    geom_bar(aes(weight=n)) +
    xlab("Year")
  
#we can also plot the distribution of area by year
ggplot(area, aes(Area_ha)) +
  geom_density() +
  facet_grid(Year ~.) +
  xlab("Area (ha)")

```

Definding patches by Year as class, we can think about things we might want to compare between runs:

- number of fires: overall use `lsp_l_np`, by year use `lsp_c_np`)
- max and median fire size: no functions, so calculate overall (lsp) and by year (class) from `lsp_p_area`
- total fire area: overall use `lsp_l_ta`, by year use `lsp_c_ca`

We can also think about [available shape metrics](https://r-spatialecology.github.io/landscapemetrics/reference/index.html#section-patch-level-metrics) (later). 

However, this approach does not allow us to calculate metrics for the continuous measures from LandTrendr. To do this we can use the `Year` layer from the original data, plus a layer with unique patch IDs created from from `get_patches()`:

```{r}
mag <- sdat.utm[["Magnitude"]]

#get patches
p <- get_patches(sdat.utm[["Year"]])
plot(stack(p))
```

Remember `get_patches()` returns a raster for each year, restarting patch numbering each year. So we need to loop to create unique patch ids across ALL years:

```{r}
patchID <- p[[1]] #initially patchID is the first layer of the stack

#loop to create unique patch ids across ALL years
for(i in 2:length(p)){

  #starting at i = 2, means val_base initially is calculated for pc
  val_base  = cellStats(patchID,stat='max',na.rm=TRUE)
  
  #update pc, adding the max value of the previous layer to all cell values
  patchID <- cover(patchID, calc(p[[i]], function(x) x + val_base))
}
```

Now create a stack of rasters containing original data, year-patches and unique patchID 

```{r}
mag.s <- stack(mag, sdat.utm[["Year"]], patchID)
names(mag.s[[3]]) <- "Patch"

plot(mag.s)

```

With this we can now calculate (zonal) statistics of magnitude by year:
```{r}

#calculate summary of magnitudes by year
zs <- zonal(mag.s[["Magnitude"]], mag.s[["Year"]], fun=sum)
zm <- zonal(mag.s[["Magnitude"]], mag.s[["Year"]], fun=median)
zsd <- zonal(mag.s[["Magnitude"]], mag.s[["Year"]], fun=sd)

mag.yr <- left_join(as.data.frame(zs), as.data.frame(zm), by = "zone")
mag.yr <- left_join(mag.yr, as.data.frame(zsd), by = "zone")
names(mag.yr) <- c("Patch", "Sum", "Median", "SD")

mag.yr

```

And by patch:
```{r}

#calculate summary of magnitudes by patch
zs <- zonal(mag.s[["Magnitude"]], mag.s[["Patch"]], fun=sum)
zm <- zonal(mag.s[["Magnitude"]], mag.s[["Patch"]], fun=median)
zsd <- zonal(mag.s[["Magnitude"]], mag.s[["Patch"]], fun=sd)

mag.patch <- left_join(as.data.frame(zs), as.data.frame(zm), by = "zone")
mag.patch <- left_join(mag.patch, as.data.frame(zsd), by = "zone")
names(mag.patch) <- c("Patch", "Sum", "Median", "SD")

mag.patch
```

It might also be useful to consider the Year in which each patch burned, so attach year to individual patches:
```{r}

#create table to cover all patchIDs (with years) to join to the mag.patch table... 
yr_ls <- unique(mag.s[["Year"]])
ids <- unique(mag.s[["Patch"]])

yr_counts <- c()
for(i in seq_along(p)){
   yr_counts <- c(yr_counts, cellStats(p[[i]],stat='max',na.rm=TRUE))
}

yrs <-cbind(yr_ls, yr_counts)
yrs <- yrs[rep(seq_len(nrow(yrs)),yrs[,2]),1:2] #expand the table
yrs <- as.data.frame(cbind(yrs, ids))

##now join to the original table 
mag.patch <- left_join(mag.patch, yrs, by = c("Patch" = "ids"))
mag.patch <- mag.patch %>%
  select(-yr_counts) %>%
  rename(Year = yr_ls)

mag.patch

```

So now for these continuous variables we could examine the following between runs (using magnitude as an example):

- total (all years) and annual magnitude between runs
- max/median magnitude (total and per year) between runs
- distribution (e.g. SD) of patch magnitude for a given year (or all years) 


