---
title: "LandTrendr-Paramo - Global Fire Atlas"
author: "James Millington"
date: "2021-07-30"
output: 
  github_document:
    toc: true
---


```{r warning=F, message=F}
library(raster)
library(sf)
library(tidyverse)
```


Load paramo boundary shapefile
```{r results = 'hide', warning=F, message=F}
GISpath <- "Data/Fire GIS Files/" 
buf <- st_read(paste0(GISpath,"500mbufferOfComplejos/GroundTruthingBuffer500m.shp"))

#set crs of shp to same as for cci raster
buf <- st_transform(buf, "+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs")
```

Load LTR dummy raster (for extent and resolution) and set projection)
```{r results = 'hide', warning=F, message=F}
#reproject takes a while so do once, write to file, then in future read 
#ltr_res <- raster("Data/NBRanalysis/LTR_JanApr_95.tif")
#ltr_res <- projectRaster(ltr_res, crs="+proj=longlat +datum=WGS84 +no_defs")
#writeRaster(ltr_res, "Data/NBRanalysis/ltr_res.grd")
ltr_res <- raster("Data/NBRanalysis/LTR_JanApr_1.tif")
```

Load DEM, set extent and resolution to LTR, mask to >2000m
```{r results = 'hide', warning=F, message=F}
#dem <- raster('Data/SRTM/srtm_22_11/srtm_22_11.tif')
#dem <- projectRaster(dem, crs="+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs", method='ngb')
#dem <- crop(dem, extent(ltr_res))
#dem <-resample(dem, ltr_res)
#writeRaster(dem, 'Data/SRTM/srtm_ltr.tif')

dem<- raster('Data/SRTM/srtm_ltr.tif')
elevize <- function(x) { ifelse(x < 2000, NA, x) }
dem2000 <- calc(dem, elevize)
```

Subset global fire atlas shapefiles to records within Paramo 2 degree box (only run once, then use output data below).
This code adds the (minimum) elevation of the point/polygon from DEM to the output data (can be used to subset by elevation later) 
```{r eval=F}
paramo_2deg_ext <- extent(-74, -72, 5, 7)  #this might be better as extent(ltr_res)

years <- 2011:2016

for(yr in years){
  point <- st_read(paste0("Data/MODIS/Global_Fire_Atlas/Global_fire_atlas_V1_ignitions_",yr,"/Global_fire_atlas_V1_ignitions_",yr,".shp"))
  point <- st_crop(point, paramo_2deg_ext)
  
  perim  <- st_read(paste0("Data/MODIS/Global_Fire_Atlas/Global_fire_atlas_V1_perimeter_",yr,"/Global_fire_atlas_V1_perimeter_",yr,".shp"))
  perim <- st_crop(perim, paramo_2deg_ext)
  
  point_dem <- extract(dem, point, sp=TRUE)
  perim_dem <- extract(dem, perim, fun=min, sp=TRUE)
  
  st_write(st_as_sf(point_dem), paste0("Data/MODIS/Global_Fire_Atlas/",yr,"/GFA_ignitions_elev_",yr,".shp"))
  st_write(st_as_sf(perim_dem), paste0("Data/MODIS/Global_Fire_Atlas/",yr,"/GFA_perimeter_elev_",yr,".shp"))
}
```

Read GFA polygons
Convert to raster with extent/resolution of LTR
  - set cell values to year of data (else zero)

```{r}

#gfa16 <- st_read(paste0("Data/MODIS/Global_Fire_Atlas/2016/GFA_perimeter_elev_2016.shp"))
#gfa16 <- rasterize(gfa16, ltr_res, field=2016, background=0)

gfa_stk <- stack()
years <- 2012:2016
for(yr in years){
  gfa  <- st_read(paste0("Data/MODIS/Global_Fire_Atlas/",yr,"/GFA_perimeter_elev_",yr,".shp"))
  
  gfa <- st_transform(gfa, "+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs")
  gfa_ras <- rasterize(gfa, ltr_res, field=yr, background=0)
  
  gfa_stk <- stack(gfa_stk, gfa_ras)
}

```


Stack rasters produced by max (work with latest fire)
  - Any cells with value >0 have fire
  - set 0s to NA
```{r}

gfa_stk_elev2000 <- mask(gfa_stk, dem2000)

gfa_max <- overlay(gfa_stk_elev2000, fun=max)
gfa_sum <- overlay(gfa_stk_elev2000, fun=sum)

NAize <- function(x) { ifelse(x == 0, NA, x) }
gfa_max <- calc(gfa_max, NAize)
gfa_sum <- calc(gfa_sum, NAize)

writeRaster(gfa_max, "Data/NBRanalysis/gfa_max_utm.grd", overwrite=TRUE)
writeRaster(gfa_sum, "Data/NBRanalysis/gfa_sum_utm.grd", overwrite=TRUE)
```

```{r}
gfa_max <- raster("Data/NBRanalysis/gfa_max_utm.grd")
gfa_sum <- raster("Data/NBRanalysis/gfa_sum_utm.grd")

```

For cells with MODIS fire, calc proportion of LTR pixels that are also fire:
[what if all are 100% covered? then need to think about how many of cells _surrounding_ the MODIS fire is 'burned' by LTR]

Load landsat raster  [change this to use SRTM DEM to decide study area]
```{r}

gfa_max <- raster("Data/NBRanalysis/gfa_max_utm.grd")

#scenario <- 1
#grp <- "JanApr"
scenario_list <- seq(from=130,to=144,by=1)
#scenario_list <- c(95,2)

for(scenario in scenario_list){
  
  for(grp in c("JanApr","AllYear")){

    print(paste0(grp,"_",scenario))
    
    #MEMORY MANAGEMENT see https://r-forge.r-project.org/forum/forum.php?thread_id=30946&forum_id=995&group_id=302
    raster_tmp_dir <- "/r_raster_tmp"  ## define the name of a temp directory where raster tmp files will be stored
    dir.create(raster_tmp_dir, showWarnings = F, recursive = T)  ## create the directory
    rasterOptions(tmpdir = raster_tmp_dir)  ## set raster options
    
    
    dat <- stack(paste0("Data/NBRanalysis/LTR_",grp,"_",scenario,".tif"))
    dat <- dat[[1]]  #layer 1 is year
    

    print(paste0("dat and gfa_max rasters are comparable: ",compareRaster(gfa_max,dat)))
    
    matches <- gfa_max == dat
    ct <- crosstab(matches,gfa_max)  #misses are 0s, hits are 1s
    
    #get false alarms [count(dat) - (hits + misses)]
    hm <- colSums(ct) #hits+misses
    fdat <- freq(dat) #counts
    fdat <- fdat[2:6,2] #get counts for 2012:2016
    
    fa <- fdat - hm     #false alarms
    ct <- rbind(ct,fa)  #bind to table
    row.names(ct) <- c("misses","hits","F-alarms") #set row.names
    
    write.csv(ct, paste0("Data/MODIS/Global_Fire_Atlas/PixelOverlap_",grp,"_",scenario,".csv"))
    
    #remember to delete the temp raster later!  for example
    unlink(raster_tmp_dir, recursive = T, force = T)

  }
}

```

Results are not good. Need to set up code to run all NBR files.... 

Kris suggests checking overlap between MODIS and year+1 for LTR  [also check overlap with GABAM... and GABAM vs Global Fire Atlas]

```{r}


#scenario <- 1
#grp <- "JanApr"
#scenario_list <- c(1,95)

scenario_list <- seq(from=1,to=144,by=1)

counts_tbl <- tibble(
  year=character(),
  pset=numeric(),
  grp=character(),
  hits=numeric(),
  misses=numeric(),
  falarms=numeric()
)

props_tbl <- tibble(
  year=character(),
  pset=numeric(),
  grp=character(),
  propHM=numeric(),
  propHF=numeric()
)

for(scenario in scenario_list){
  
  for(grp in c("JanApr","AllYear")){
    
    po <- read_csv(paste0("Data/MODIS/Global_Fire_Atlas/PixelOverlap_",grp,"_",scenario,".csv"), 
                   col_select=-1, 
                   col_types=cols(.default="i"))

    misses <- po %>% 
      slice(1) %>%
      pivot_longer(cols=everything(), names_to="year",values_to="misses")
    
    hits <- po %>% 
      slice(2) %>%
      pivot_longer(cols=everything(), names_to="year",values_to="hits")
      
    falarm <- po %>% 
      slice(3) %>%
      pivot_longer(cols=everything(), names_to="year",values_to="falarms")
    
    counts <- inner_join(hits,misses,by="year")
    counts <- inner_join(counts,falarm,by="year")
    counts <- mutate(counts, pset=scenario,grp=grp)
    counts_tbl <- bind_rows(counts_tbl, counts)
    
    #proportion of hits to misses
    propHM <- round(po[2,]/po[1,],5)
    propHM <-pivot_longer(propHM, cols=everything(), names_to="year",values_to="propHM")
    
    #proportion of hits to false alarms
    propHF <- round(po[2,]/po[3,],5)
    propHF <-pivot_longer(propHF, cols=everything(), names_to="year",values_to="propHF")
    
    props <- inner_join(propHM,propHF,by="year")
    props <- mutate(props, pset=scenario,grp=grp)
    props_tbl <- bind_rows(props_tbl, props)
  }
}

write_csv(counts_tbl, "Data/MODIS/Global_Fire_Atlas/PixelOverlap_Counts.csv")
write_csv(props_tbl, "Data/MODIS/Global_Fire_Atlas/PixelOverlap_Props.csv")


```

```{r}

plotCounts <- function(.dat, .var){

  #.dat=countsDat
  #.var="misses"
  
  maxcount <- .dat %>%
    select(year:grp, {{.var}}) %>%
    group_by(pset, grp) %>%
    summarise(counts = sum(.data[[.var]])) 
    
  maxcount <- ceiling(max(maxcount$counts))
  
  #for loop across parameter sets (to split plots for legibility)
  Pset_filter <- data.frame(rbind(c(0,24),c(24,48),c(48,72),c(72,96),c(96,120),c(120,144)))
  
  #row=1
  #loop over pset groups and plot
  for(row in 1:length(Pset_filter[,1])){
    
    lower <- Pset_filter[row,1]
    upper <- Pset_filter[row,2]
  
    plotDat <- .dat %>%
      select(year:grp, {{.var}}) %>%
      filter(pset > lower & pset <= upper) %>%
      mutate(pset = formatC(pset, width = 3, format = "d", flag = "0")) %>%  #pad with preceeding 0s so that plots margins are identical
      mutate(pset = factor(pset)) %>%
      mutate(year = factor(year))
  
    p <- ggplot(plotDat, aes(fill=year, y=.data[[.var]], x=grp)) + 
      geom_bar(position="stack", stat="identity") +
      facet_grid(.~pset) +
      ylim(0,maxcount) +
      labs(title=.var) + 
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
    print(p)
  }
}

```

```{r}


countsDat <- read_csv(paste0("Data/MODIS/Global_Fire_Atlas/PixelOverlap_Counts.csv"))
#Mall <- read_csv(paste0("Data/MODIS/Global_Fire_Atlas/PixelOverlap_Misses.csv"))
#Fall <- read_csv(paste0("Data/MODIS/Global_Fire_Atlas/PixelOverlap_FalseAlarms.csv"))

pdf("Data/MODIS/Global_Fire_Atlas/Hits_PixelOverlap.pdf", width=10, height=10)
plotCounts(countsDat, "hits")
dev.off()

pdf("Data/MODIS/Global_Fire_Atlas/Misses_PixelOverlap.pdf", width=10, height=10)
plotCounts(countsDat, "misses")
dev.off()

pdf("Data/MODIS/Global_Fire_Atlas/FAlarms_PixelOverlap.pdf", width=10, height=10)
plotCounts(countsDat, "falarms")
dev.off()

```

Plot:
- proportion of hits to misses
- proportion of hits to false alarms


```{r}

plotPropHM <- function(.dat){

  #for loop across parameter sets (to split plots for legibility)
  Pset_filter <- data.frame(rbind(c(0,24),c(24,48),c(48,72),c(72,96),c(96,120),c(120,144)))
  row <- 1
  #loop over pset groups and plot
  for(row in 1:length(Pset_filter[,1])){
    
    lower <- Pset_filter[row,1]
    upper <- Pset_filter[row,2]
  
    plotDat <- .dat %>%
      filter(pset > lower & pset <= upper) %>%
      mutate(pset = formatC(pset, width = 3, format = "d", flag = "0")) %>%  #pad with preceeding 0s so that plots margins are identical
      mutate(pset = factor(pset)) %>%
      mutate(year = factor(year)) 
  
    p <- ggplot(plotDat, aes(fill=year, y=count, x=grp)) + 
      geom_bar(position="stack", stat="identity") +
      facet_grid(.~pset) +
      ylim(0,maxcount) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
    print(p)
  }
}

```

```{r}
matches1 <- gfa_max == dat
#gfa16 <- calc(gfa16, NAize)

crosstab(matches1,gfa_max)
```

```{r}

dat1 <- mask(dat1, dem2000)

dat16 <- calc(dat1, NAize)

```

```{r}
perim  <- st_read(paste0("Data/MODIS/Global_Fire_Atlas/2016/GFA_perimeter_elev_2016.shp"))

plot(dat16, col='red')
zoom(dat16)

plot(perim, add=T)
freq(dat16)
```


Create polygon of our max study extent
```{r}
bb <- bbox(dat1)

#following from https://gis.stackexchange.com/a/403979
bb_pol = st_polygon(
  list(
    cbind(
      bb[1,][c(1,2,2,1,1)], 
      bb[2,][c(1,1,2,2,1)])
  )
)

#set projection to original raster (dat1)
bb_pol = st_sfc(bb_pol, crs="+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
#reproject to GFA raster
bb_pol <- st_transform(bb_pol, "+proj=longlat +datum=WGS84 +no_defs")
```

Clip points to bounding box
```{r}

#i16_p <- st_intersection(i16, bb_pol)
#plot(i16_p)

p16_p <- st_intersection(p16, bb_pol)
plot(p16_p)

```


```{r}

plot(st_geometry(p16_p),col='black',pch=3)
plot(st_geometry(buf),border='red',add=T)

```

Next Steps
- Combined all GFA years?
- Filter LTR data by size (min 21 ha)?

```{r}
#function to set 0 to NA in raster
NAize <- function(x) { ifelse(x == 0, NA, x) }

#function to set 0 to NA in stack
myStack_NAize <- function(myStack){
  yr <- calc(myStack[[1]],NAize)
  myStack <- mask(myStack, yr)
  return(myStack)
}
```

```{r}

stk <- stack('/home/james/wdResearch/LandTrendr-Paramo/Data/NBRanalysis/LTR_JanApr_96.tif')
stk <- myStack_NAize(stk)

ras <- stk[[1]]
ras <- projectRaster(ras, crs="+proj=longlat +datum=WGS84 +no_defs")
crs(ras)
summary(ras)
plot(stk)
```

```{r}

plot(ras)
plot(st_geometry(p16_p),col='black',pch=3,add=T)
plot(st_geometry(buf),border='red',add=T)

```