---
title: "LandTrendr-Paramo - comparisons"
author: "James Millington"
date: "2021-09-09"
output: 
  github_document:
    toc: true
---


```{r warning=F, message=F}
library(raster)
library(sf)
library(tidyverse)
library(rasterVis)
```

unmasked maps
```{r}

GISpath <- "Data/Fire GIS Files/" 
buf <- st_read(paste0(GISpath,"500mbufferOfComplejos/GroundTruthingBuffer500m.shp"))
buf <- st_transform(buf, "+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs")

gfa_max <- raster("Data/NBRanalysis/gfa_max_utm.grd")
gabam_max <- raster("Data/NBRanalysis/gabam_max_utm.grd")

ltr127ay <- raster("Data/NBRanalysis/LTR_AllYear_127.tif")
NAize <- function(x) { ifelse(x == 0, NA, x) }
ltr127ay <- calc(ltr127ay, NAize)

plot(gfa_max)
plot(st_geometry(buf), add=T)

plot(gabam_max)
plot(st_geometry(buf), add=T)

plot(ltr127ay)
plot(st_geometry(buf), add=T)

```
masked maps
```{r}
GISpath <- "Data/Fire GIS Files/" 
buf <- st_read(paste0(GISpath,"500mbufferOfComplejos/GroundTruthingBuffer500m.shp"))
buf <- st_transform(buf, "+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs")

gfa_max_masked <- raster("Data/NBRanalysis/gfa_max_masked.grd")
gabam_max_masked <- raster("Data/NBRanalysis/gabam_max_masked.grd")

mymask <- raster("/home/james/wdResearch/LandTrendr-Paramo/Data/GIS/GroundTruthingBuffer500m_mask_trim.tif")


ltr126ay <- raster("Data/NBRanalysis/LTR_AllYear_126.tif")

ltr126ay <- crop(ltr126ay, mymask)
ltr126ay <- mask(ltr126ay, mymask) 

NAize <- function(x) { ifelse(x == 0, NA, x) }
ltr126ay <- calc(ltr126ay, NAize)

maxize <- function(x) { ifelse(x > 2016, NA, x) }
ltr_masked <- calc(ltr126ay, maxize)

plot(gfa_max_masked)
plot(st_geometry(buf), add=T)

plot(gabam_max_masked)
plot(st_geometry(buf), add=T)

plot(ltr_masked)
plot(st_geometry(buf), add=T)

```


```{r}

mymask <- raster("/home/james/wdResearch/LandTrendr-Paramo/Data/GIS/GroundTruthingBuffer500m_mask_trim.tif")

maxize <- function(x) { ifelse(x > 2016, NA, x) }
ltr_max<- calc(ltr126ay, maxize)
ltr_crop <- crop(ltr_max, mymask)
ltr_masked <- mask(ltr_crop, mymask) 

gfa_max_crop <- crop(gfa_max, mymask)
gfa_max_masked <- mask(gfa_max_crop, mymask) 

gabam_max_crop <- crop(gabam_max, mymask)
gabam_max_masked <- mask(gabam_max_crop, mymask) 

plot(gfa_max_masked)
plot(st_geometry(buf), add=T)

plot(gabam_max_masked)
plot(st_geometry(buf), add=T)

plot(ltr_masked)
plot(st_geometry(buf), add=T)

```

```{r}

oneize <- function(x) { ifelse(x > 0, 1, NA) }
ltr1s<- calc(ltr_masked, oneize)
ltr1s_proj <- projectRaster(from = ltr1s, crs = CRS("+init=epsg:3857"),method='ngb')

gabam1s<- calc(gabam_max_masked, oneize)
gabam_proj <- projectRaster(from = gabam1s, crs = CRS("+init=epsg:3857"),method='ngb')

gfa1s<- calc(gfa_max_masked, oneize)
gfa_proj <- projectRaster(from = gfa1s, crs = CRS("+init=epsg:3857"),method='ngb')


buf_proj <- st_transform(buf, crs=3857)
```

```{r}
redPal <- RColorBrewer::brewer.pal('Set1', n=1)
redTheme <- rasterTheme(region = c('#FF0000'))

ltrlp <- levelplot(ltr1s_proj, margin=F, par.settings=redTheme, colorkey=F)
ltrlp + latticeExtra::layer(sp.polygons(as_Spatial(from = st_geometry(buf_proj)),col="black"))

gabamlp <- levelplot(gabam_proj, margin=F, par.settings=redTheme, colorkey=F)
gabamlp + latticeExtra::layer(sp.polygons(as_Spatial(from = st_geometry(buf_proj)),col="black"))

gfslp <- levelplot(gfa_proj, margin=F, par.settings=redTheme, colorkey=F)
gfslp + latticeExtra::layer(sp.polygons(as_Spatial(from = st_geometry(buf_proj)),col="black"))
```


```{r}
gfa_freq <- freq(gfa_max_masked)
gabam_freq <- freq(gabam_max_masked)
ltr_freq <- freq(ltr_masked)

gfa_freq <- rbind(gfa_freq, c(2012,0))

all_freq <- left_join(as.data.frame(gfa_freq), as.data.frame(gabam_freq), by = "value", suffix=c(".gfa",".gabam"))
all_freq <- left_join(all_freq, as.data.frame(ltr_freq), by = "value")

  
all_ha <- all_freq %>%
  mutate(gfa = count.gfa * 0.09) %>% 
  mutate(gabam = count.gabam * 0.09) %>%
  mutate(ltr = count * 0.09) %>%
  rename(year = value) %>%
  pivot_longer(cols=c("gfa","gabam", "ltr"),
  #pivot_longer(cols=c("gfa","gabam"),
               names_to="sensor",
               values_to="ha") %>%
  select(-starts_with("count")) %>%
  drop_na()

ggplot(all_ha, aes(x=year, y=ha, fill=sensor)) +
  geom_bar(stat="identity",position="dodge") +
  scale_fill_viridis_d("Product") +
  xlab("Year") +
  ylab("Burned Area (ha)") +
  ggtitle("Burned Area by Product and Year")

```

Calculate burned area for all LTR psets 
```{r}

scenario_list <- seq(from=1,to=144,by=1)
#scenario_list <- c(95,2)

ha_tbl <- tibble(
  pset=numeric(),
  grp=character(),
  '2012'=numeric(),
  '2013'=numeric(),
  '2014'=numeric(),
  '2015'=numeric(),
  '2016'=numeric()
)

for(scenario in scenario_list){
  
  for(grp in c("JanApr","AllYear")){
    
    #scenario <- 141
    #grp <- "AllYear"
    dat <- read_csv(paste0("/home/james/wdResearch/LandTrendr-Paramo/Data/GABAM/PixelOverlap_",grp,"_",scenario,"gabam_masked.csv"))
    
    dats <- dat %>%
      summarise_if(is.numeric, ~ sum(.x) * 0.09) %>%
      #across(.cols = everything(), ~ .x * 0.09) %>%
      mutate(pset=scenario, grp=grp)
      
    ha_tbl <- bind_rows(ha_tbl, dats)
    
  }
}

write_csv(ha_tbl,"Data/NBRanalysis/LTRha_masked.csv")

```

Plot LTR burned area using output of code block above 
```{r}

ltr_ha <- read_csv("Data/NBRanalysis/LTRha.csv")

ltr_ha_long <- ltr_ha %>%
  filter(grp=="AllYear") %>%
  select(-grp) %>%
  pivot_longer(`2012`:`2016`, values_to = "ha", names_to="year")

ggplot(ltr_ha_long, aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge")

```

Plot LTR burned area using summary of counts
```{r}

ltr_counts <- read_csv("Data/GABAM/PixelOverlap_Counts-gabam_masked.csv")

ltr_ha_long <- ltr_counts %>%
  rowwise() %>%
  mutate(ha = sum(hits, falarms) * 0.09) %>%
  mutate(year = as.factor(year)) %>%
  select(-hits, -misses, -falarms)

ggplot(ltr_ha_long, aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge")
         

```


plot specific combinations of LTR psets 
``` {r}

selection <- c(79,80,87,88,95,96,103,104,111,112,119,120,127,128,135,136,143,144) 
mag400 <- c(80,88,96,104,112,120,128,136,144)  #magnitude = 400
mag350 <- mag400 - 1   #magnitude = 350
mag300 <- mag350 - 1   #magniture = 300
mag250 <- mag300 - 1   #magniture = 250


p005 <- c(79, 103, 127)  #mag = 350, pval = 0.05

ltr_ha_long %>%
  filter(pset %in% mag400) %>%
  ggplot( aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge") +
  facet_grid(.~grp) +
  ggtitle("mag400")

ltr_ha_long %>%
  filter(pset %in% mag350) %>%
  ggplot( aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge") +
  facet_grid(.~grp) +
  ggtitle("mag350")

ltr_ha_long %>%
  filter(pset %in% mag300) %>%
  ggplot( aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge") +
  facet_grid(.~grp) +
  ggtitle("mag300")

ltr_ha_long %>%
  filter(pset %in% mag250) %>%
  ggplot( aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge") +
  facet_grid(.~grp) +
  ggtitle("mag250")

ltr_ha_long %>%
  filter(pset %in% p005) %>%
  ggplot( aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge") +
  facet_grid(.~grp) +
  ggtitle("p005")

```

```{r}

ltr_ha_long_grp <- ltr_ha %>%
  #filter(grp=="AllYear") %>%
  #select(-grp) %>%
  pivot_longer(`2012`:`2016`, values_to = "ha", names_to="year")

ltr_ha_long_grp %>%
  filter(pset %in% mag350) %>%
  ggplot( aes(x=pset,y=ha,fill=year)) +
         geom_bar(stat="identity",position="dodge") +
  facet_grid(grp~.) +
  ggtitle("LTR (Mag = 350)")


```

Compare spatial distribution of the psets above - does that help to choose? I guess the spatial distribution among LTR will be very similar?  

Check by comparing pixel ids? 

Or convert to binary (fire, no fire) and sum (assumes years will be the same)

If all very similar then pick one from odds above and one from evens (e.g. 143 and 144) and compare each to GFA and GABAM respectively.

``` {r}
#need to run this code on windows machine... 

binarize <- function(x) { ifelse(x > 0, 1, NA) }

for(scenario in selection){
  
  for(grp in c("JanApr","AllYear")){

    print(paste0(grp,"_",scenario))
    
    #MEMORY MANAGEMENT see https://r-forge.r-project.org/forum/forum.php?thread_id=30946&forum_id=995&group_id=302
    raster_tmp_dir <- "/r_raster_tmp"  ## define the name of a temp directory where raster tmp files will be stored
    dir.create(raster_tmp_dir, showWarnings = F, recursive = T)  ## create the directory
    rasterOptions(tmpdir = raster_tmp_dir)  ## set raster options
    
    
    dat <- stack(paste0("Data/NBRanalysis/LTR_",grp,"_",scenario,".tif"))
    dat <- dat[[1]]  #layer 1 is year
    
    bindat <- calc(dat, binarize)
    
    writeRaster(bindat, paste0("Data/NBRanalysis/LTR_",grp,"_",scenario,"_binary.tif"))
  }
}


```
